---
title: "PCA and MDS of American Bankruptcy between 1980 and 2000 "
output: html_document
authors : Jason Ching Yuen Siu, Liguo Bao, Wanxin Liu, Zhixiang Yang
---




```{r include=FALSE}

knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 5, warning = F)
library(tidyverse)
library(naniar)
library(Hmisc)
library(broom)
library(fastDummies)
library(plotly)
library(viridis)

# MDS 
remotes::install_github("AckerDWM/gg3D")
library(gg3D)  # 这是一个3d数据包
library(rgl)
library(gg3D)
library(vegan)
library(plotly)
library(caret)

#PCA 2D
library(broom)
library(dummies)
library(tidyr)

#PCA3D
library(plot3D)
library(plot3Drgl)
Bankruptcy <- readRDS('Bankruptcy.RDS')
```

======================================== Preliminary Analysis ======================================== 


View NA
```{r}
png(file="PA_Missingness.png", res=300, width=1600, height=1000)
vis_miss(Bankruptcy, sort=T) +
  ggtitle("Missingness of the Bankruptcy dataset")+
  theme(axis.text.x=element_text( angle = 90,size=8))
```

- Only 2.1% of the data has NA values, which is somewhat clean.
- However, since EmplUnion has 30.96 % of NA values, this feature is eliminated.
- FirmEnd and DENYOther brings little values in this analysis, therefore they are also eliminated.

Exclude non-needed attributes
```{r}
Bankruptcy <- Bankruptcy %>% select(-EmplUnion,-FirmEnd)
```

- Since the economy is analysed by quarter usually, therefore, Quarter is sorted.
```{r}
# add an index to the dataset
Bankruptcy <- Bankruptcy %>% mutate(ID = row_number())
# make a quarter
Bankruptcy <- Bankruptcy %>% mutate(Quarter = ceiling(MonthFiled / 3))
```
- There are 55 industries. It will be difficult to conduct an anlysis with so many categorical values; 
hence they are categorised into different sectors.
```{r}
# Categorise sector
agri <- tibble(Industry = unique(Bankruptcy$SICMajGroup) )%>%
  filter(row_number()==53) 
financial <- tibble(Industry = unique(Bankruptcy$SICMajGroup) )%>%
  filter(row_number() %in% c(50,17,30)) 
property <- tibble(Industry = unique(Bankruptcy$SICMajGroup) )%>%
  filter(row_number() %in% c(27,33,34,37,29,42)) 
service <- tibble(Industry = unique(Bankruptcy$SICMajGroup) )%>%
  filter(row_number() %in% c(22,28,47,51,8,24,54,55,11,38,20,52))

industrial <- tibble(Industry = unique(Bankruptcy$SICMajGroup) )%>%
  filter(row_number() %in% c(4,6,9,1,21,39,41,13,2,35,19,3,45,14))

consumption <- tibble(Industry = unique(Bankruptcy$SICMajGroup) )%>%
  filter(row_number() %in% c(43,44,40,32,23,31,15,46,25,18,16,5,7,36,12))

resource <- tibble(Industry = unique(Bankruptcy$SICMajGroup) )%>%
  filter(row_number() %in% c(10,26,48,49))

b <- data.frame()
assign_sector <- function(industry,sector){
  for(i in 1:length(industry)){
  a <- Bankruptcy %>% filter(SICMajGroup %in% industry$Industry)  %>% mutate(Sector = sector)
  b <-   rbind(a,b)
  }
  return(b)
}

df1 <- assign_sector(financial,"Financial")
df2 <- assign_sector(property, "Property")
df3 <- assign_sector(service, "Service")
df4 <- assign_sector(industrial, "Industrial")
df5 <- assign_sector(consumption, "Consumption")
df6 <- assign_sector(agri,"Agriculture")
df7 <- assign_sector(resource,"Resource")

Bankruptcy <-  rbind(df1,df2,df3,df4,df5,df6, df7)
#Denyother enumerate
Bankruptcy$DENYOther_id <- Bankruptcy$DENYOther %>% as.factor() %>% as.numeric()
#filter df for PA
df_for_desc_stat<-  rbind(df1,df2,df3,df4,df5,df6, df7)
#filter df for MDS
new_var<-Bankruptcy[,c("Name","CPI","DaysIn","Ebit","Employees","FilingRate","GDP","HeadCityPop","HeadCourtCityToDE","Liab","Sales","HeadStAtFiling","Sector","YearFiled","Quarter","CityFiled","DENYOther","MonthFiled","ID","DENYOther_id")]%>%drop_na()
#filter df for PCA3D
Bankruptcy%>% 
  filter(ID != 58) %>% 
  column_to_rownames('ID') %>% 
  select_if(is.numeric) %>%
  select(-Quarter,-DENYOther_id) %>% 
  na.omit() %>% 
  filter()-> pca_data

```
Categorise variables
```{r factor}

Bankruptcy$CityFiled <- as.factor(Bankruptcy$CityFiled) %>% as.numeric()
Bankruptcy$HeadStAtFiling <- as.factor(Bankruptcy$HeadStAtFiling) %>% as.numeric()
Bankruptcy$SICMajGroup <- as.factor(Bankruptcy$SICMajGroup) %>% as.numeric()
Bankruptcy$Sector <- as.factor(Bankruptcy$Sector)%>% as.numeric()
Bankruptcy$Assets <- as.factor(Bankruptcy$Assets) %>% as.numeric()
Bankruptcy$DaysIn <- as.factor(Bankruptcy$DaysIn) %>% as.numeric()
Bankruptcy$Employees <- as.factor(Bankruptcy$Employees) %>% as.numeric()
Bankruptcy$FilingRate <- as.factor(Bankruptcy$FilingRate) %>% as.numeric()
Bankruptcy$HeadCourtCityToDE <- as.factor(Bankruptcy$HeadCourtCityToDE) %>% as.numeric()
Bankruptcy$MonthFiled <- as.factor(Bankruptcy$MonthFiled) %>% as.numeric()
Bankruptcy$YearFiled <- as.factor(Bankruptcy$YearFiled) %>% as.numeric()


```




```{r select-numeric}
bc_numeric <- Bankruptcy %>% select_if(is.numeric) 
```

```{r boxplot}
company <- c("Assets","Ebit","Employees","Liab")
general <- c("FilingRate","Quarter","CityFiled")

#Company
boxp<-ggplot(
  gather(
    #remove unneeded columns
    bc_numeric %>% select(company))
  , aes(value)) + 
    geom_boxplot(bins = 10) + 
    facet_wrap(~key, scales = 'free_x')+
  theme_bw()+
    theme(axis.text.x = element_text(angle = 360, vjust = 1, hjust = 1))+
    geom_boxplot(color="red", fill="orange", alpha=0.2)+
  xlab("")+ylim(10,500)
ggplotly(boxp) 



```
```{r heatmap-macro, warning=F}

png(file="PA_heatmap.png", res=300, width=2000, height=1000)

  grp_yr <- df_for_desc_stat %>% group_by(YearFiled, Quarter) %>% 
  summarise(GDP=median(GDP), 
        CPI=median(CPI))%>% 
        arrange(desc(YearFiled)) 
grp_yr <- grp_yr %>%   pivot_longer(cols = c("CPI","GDP"), names_to = "Type", 
values_to= "Values") 
 
 mycol <- c("navy", "blue", "cyan", "lightcyan", "yellow", "red", "red4") 
        ggplot(grp_yr, aes(x= YearFiled , y=as.integer(Quarter), color = Values)) + 
        geom_tile (aes(fill=Values),colour = "white" ) + 
        scale_fill_gradientn(colours = mycol)+ 
        facet_grid(~Type)+ 
        theme_linedraw()+ 
        theme(axis.text.x = element_text( color="BLACK", angle=90),
              axis.text.y = element_text( color="BLACK", angle=90))+
          labs( 
        title = "Time-Series Calendar Heatmap: GDP and CPI", 
        x = "Year", 
        y = "Quarter" 
        )  +theme(text=element_text(family="Times New Roman", face="bold", size=12)) 
        
```


```{r Liab, warning=F}
#Which sector have the most liability?

png(file="PA_sector_bar.png", res = 300,width=2000, height=1000)
ggplot(df_for_desc_stat, aes(fct_reorder(Sector, Liab), y =Liab)) +
  geom_col(fill="dark blue")+
  theme_classic()+    
  theme(      
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    text=element_text(family="Times New Roman", face="bold", size=12)
    )+
  ggtitle("The liability by Sector")+
  ylab("Liability")+xlab("Sector")


# which state have the most liability?
png(file="PA_states_bar.png", res = 300,width=2000, height=1000)
ggplot(df_for_desc_stat, aes(x=fct_reorder(HeadStAtFiling,Liab), y=Liab )) + geom_col(fill="dark blue")+
  theme_classic()+    
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    text=element_text(family="Times New Roman", face="bold", size=12))+
  ggtitle("The liability by State")+
  ylab("Liability")+xlab("State")

#Which year has the highest number of bankruptcy ?
png(file="PA_year_bar.png", res = 300,width=2000, height=1000)

ggplot(df_for_desc_stat, aes(x=factor(YearFiled), y=Liab, fill =factor(Quarter) )) + 
    geom_bar(position="stack", stat="identity")+
    theme_classic()+    
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    text=element_text(family="Times New Roman", face="bold", size=12))+
  ggtitle("The liability by Year")+
  ylab("Liability")+xlab("Year")    +
  scale_fill_viridis(discrete = T, name="Quarter") 



png(file="PA_states_bar_DenyOther.png", res = 300,width=2000, height=1000)
df <- df_for_desc_stat %>% group_by(Sector, DENYOther) %>% tally()
ggplot(df, aes(fill=DENYOther, y=n, x=Sector)) + 
    geom_bar(position="stack", stat="identity")+theme_classic()+  scale_fill_viridis(discrete = T, name="City") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    text=element_text(family="Times New Roman", face="bold", size=12))+
  ylab("Amount")+
  ggtitle("The amount of filed bankruptcy by city")


#state = x; #bc = y; color = in

df <- df_for_desc_stat %>% group_by(Sector, HeadStAtFiling) %>% tally()
  png(file="states_bar_State.png", res = 300,width=2000, height=1000)
  ggplot(df, aes(fill=Sector, y=n, x=HeadStAtFiling)) + 
      geom_bar(position="stack", stat="identity")+theme_classic()+  scale_fill_viridis(discrete = T, name="Sector") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      text=element_text(family="Times New Roman", face="bold", size=12))+
    ylab("Amount")+xlab("Sector")+
    ggtitle("The amount of filed bankruptcy in different states by sector")


```


```{r #of File}
Bankruptcy %>%
 mutate(Quarter = factor(tools::toTitleCase(as.character(Quarter)))) %>%
   ggplot() +
    geom_histogram(aes(Quarter, fill = Quarter), stat = "count") +
   theme_minimal(base_size = 10) +
   ggtitle("Number of companies filed per quarter") +
   scale_fill_brewer(palette = "Set3") +
   xlab("Quarter ") +
   ylab("Num. Companies")->p11
 
p11
```

```{r corr-matrix}
library(corrplot)

png(file="corr.png", res=300, width=4500, height=4500)

corbank<-cor(bc_numeric %>% drop_na()) %>% as.matrix()

corrplot(corbank,
         method="color",
         type="upper",
         order="hclust",diag = FALSE,
         addCoef.col = "black",sig.level = 0.05,
         insig="blank",
         number.cex=1,
         mar=c(0,1,1,5),
         main="The correlation graph between numerical variables")

```


======================================== MDS ======================================== 


```{r df_for_MDS}


NumSec<-as.matrix(new_var["Sector"])
NumSector<-as.numeric(factor(NumSec))
NumSector<-tibble(NumSector)
new_var<-cbind(new_var,NumSector)



num_var<-new_var[,c("CPI","DaysIn","Ebit","Employees","FilingRate","GDP","HeadCityPop","HeadCourtCityToDE","Liab","NumSector","Quarter","ID","DENYOther_id")]

dropped<-num_var%>%drop_na()



## here how to adjust the margin again with more clearer display 
```




## Goodness of fit 

```{r}

delta1<-filter(num_var)%>%
   select_if(is.numeric)%>%
   scale%>%
   dist(method="manhattan")

attributes(delta1)$Labels<-filter(new_var)%>%
   pull(HeadStAtFiling)%>%
   abbreviate(6)
 ## delta1 is manhattan distance but same as Euclidean Classical method 

delta<-filter(num_var)%>%
   select_if(is.numeric)%>%
   scale%>%
   dist



 ## Euclidean distance data in delta 


attributes(delta)$Labels<-filter(new_var)%>%
   pull(CityFiled)%>%
   abbreviate(6)




mdsout1<-cmdscale(delta1,eig = TRUE) ##delta1 表示的Manhattan 距离
str(mdsout1$GOF)

mdsout<-cmdscale(delta,eig=TRUE)## delta表示Euclidean距离

str(mdsout$GOF)

eigval<-mdsout$eig 

```



## 3D Euclidean MDS 



```{r scaling}
normalize <- function(x, na.rm = TRUE) {
    return((x- min(x)) /(max(x)-min(x)))
}
```
 

```{r}

# 这个就是转换成三维坐标写的，还是classical的模型

data.dist<-dist(dropped %>% normalize()) ## euclidean distance , normalised
data.mds<-cmdscale(data.dist,k=3) ## Generate euclidean 

#plot3d(data.mds,col=dropped$NumSector)


df11<-data.frame(var1=data.mds[,1],var2=data.mds[,2],var3=data.mds[,3],Employees=dropped$Employees) ## EUCLIDEAN  datast 

d3plot<-ggplot(df11,aes(x=var1,y=var2,z=var3,color=Employees))+theme_void() +axes_3D()+stat_3D() + scale_colour_gradient(
  low = "orange",
  high = "dark blue",
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "colour"
)+ggtitle("Relationship between employees and companies")+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),text=element_text(family="Times New Roman", face="bold", size=12))

png(file="MDS_relation_bt_emp_comp.png", res=300, width=4500, height=4500)

ggplotly(d3plot)


#plot_ly(data=df11,x=var1,y=var2,z=var3,type="scatter3d", mode="markers")

```


 
##relationships between headquarters 
 
```{r}

# 这个就是转换成三维坐标写的，还是classical的模型

data.dist<-dist(dropped %>% normalize()) ## euclidean distance , normalised
data.mds<-cmdscale(data.dist,k=3) ## Generate euclidean 



df11<-data.frame(var1=data.mds[,1],var2=data.mds[,2],var3=data.mds[,3],dishead=dropped$HeadCourtCityToDE) ## EUCLIDEAN  datast 

d3plot<-ggplot(df11,aes(x=var1,y=var2,z=var3,color=dishead))+theme_void() +axes_3D()+stat_3D() + scale_colour_gradient(
  low = "orange",
  high = "dark blue",
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "colour"
)+ggtitle("Relationship between distances of headquarter and bankrupcy")

png(file="MDS_relation_bt_head_bank.png", res=300, width=4500, height=4500)
ggplotly(d3plot)


```



# Sammon mapping 
```{r SM, warning=F}
library(MASS)
dropped.sammon<-sammon(dist(dropped),k=2) ## 二维sammon的坐标在这里
# stress 
dropped.sammon$stress


scree.plot = function(d, k) {
    stresses=sammon(d, k=k)$stress
    for(i in rev(seq(k-1)))  
        stresses=append(stresses,sammon(d, k=i)$stress)
    plot(seq(k),rev(stresses), type="b", xaxp=c(1,k, k-1), ylab="Stress", xlab="Number of dimensions",main="Scree plot via sammon mapping")
}


png(file="MDS_screeplot.png", res=300, width=4500, height=4500)
scree.plot(dist(dropped),k=6)## scree plot 对于sammon来说


## sammon mapping 
df13<-data.frame(v1=dropped.sammon$points[,1],
                 v2=dropped.sammon$points[,2],
                 Sector=factor(NumSec),DENYother=factor(new_var$DENYOther)) ## 这个sammon的图建议还是把数字换成industry的名称!! 

colors <-c("Financial" = "#F51720", "Property" ="Blue" , "Industrial" = "Yellow", "Industrial"="#4C5270" , "Consumption" = "#F8D210", "Agriculture" = "#FA26A0", "Resource"  = "dark blue" )

png(file="MDS_sector.png", res=300, width=4500, height=4500)

ggplot(df13,aes(x=v1,y=v2,col=Sector))+
   geom_point(size = 4, alpha =.5)+  
   scale_shape_manual(values = c(10:16))+ 
  theme_classic() +ggtitle("The Sector of bankrupted companies")+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    text=element_text(family="Times New Roman", face="bold", size=12))

```





## relationship between year filed and the legnth of Bankrupcy 
```{r rug , warning=F}

## with the year increased , the length of bankrupcy becomes faster than previous years 
## previous year ggplot(new_var,aes(x=DaysIn,col=YearFiled))+geom_rug()+ggtitle("The length of bankrupcy by year") 

df15<-data.frame(DaysIn=new_var$DaysIn,DENYOth=new_var$DENYOther,YearFiled=factor(new_var$YearFiled))
png(file="MDS_length_BC_by_yr.png", res=300, width=4500, height=4500)
ggplot(df15,aes(x=DaysIn,y=DENYOth,col=YearFiled))+
  geom_point()+
  ggtitle("The length of bankrupcy by year")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    text=element_text(family="Times New Roman", face="bold", size=12))+  
  scale_fill_viridis(discrete = T, name="YearFiled") +theme_bw()
```




## NMDS 



```{r NMDS, warning=F}
set.seed(123456)
dropped.nmds<-metaMDS(dropped,distance="bray",autotransform = FALSE)

dropped.envfit<-envfit(dropped.nmds,dropped,permutations=999)



dropped.nmds
```


#DENYother

```{r ordiplot}
png(file="MDS_ordiplot.png", res=300, width=4500, height=4500)


ordiplot(dropped.nmds,type="n", main = "Different locations NMDS")

#ordiplot(dropped.nmds,type="n", main = "different quarters")

orditorp(dropped.nmds, display = "sites", labels = F, col = c("green", "blue","red") [as.numeric(dropped$DENYOther_id)], cex = 1)

# other green, NY blue, DE red 
legend("bottomleft", legend=c("OT", "NY", "DE"), 
       col = c("green", "blue","red"),lty=1, cex=1,box.lty=2, box.lwd=1)

```


======================================== PCA2D ======================================== 



# Principal Components Analysis and Biplot

Principal components analysis is a convenient way to merge variables through a linear combination. Since this report wants to investigate the reason of bankruptcy in different observations and more convenient views, we set ID as row names instead of certain names. After removing missing values, numerical variables are selected because categorical variables cannot be used by PCA. In addition, original variables are measured in different units, so we standardise the data by dividing by the standard deviation.

In the whole PCA analysis, the first two components just explain 49.46% variance, which is lower. However, we can firstly investigate some insights from the first two components. The first principal component is a measure of macro attributes of bankrupt companies, since it is positively correlated with common macro features of improving bankrupt rate, such as filling rate, CPI, yearfield and GDP whereas it is negatively correlated with common macro features of declining bankrupt rate, such as PrimeFiling. Similarly, the second principal components is a measure of micro attributes of bankrupt companies because it is positively correlated with common micro features of improving bankrupt rate, such as Employees, Ebit, Liab, Sales and Assets. Moreover, some loadings of variables are small and no significant effect on PC1 and PC2, so we decide to explain those variables in 3d PCA plots later.

Most observations are focused on the same position, which means that those bankruptcy companies have similar features. No.58 company is an significant outlier in the plot, and this company is called Texaco Inc. It is the world’s third-largest oil company, and the biggest firm ever to seek bankruptcy protection. Texaco was forced into bankruptcy as a result of a 1985 decision by a Houston jury that Texaco had illegally interfered with Pennzoil’s plan to acquire Los Angeles-based Getty Oil Co. The jury awarded Pennzoil damages of $10.5 billion plus interest. The PCA plot also confirms it. Texaco Inc. has the largest number of assests, liab, ebit and employees.

```{r function-for-adding-title}
add_title <-  function(sector,title_at,subtitle_at){
mytitle = "Principal Component Analysis"
mysubtitle = paste("",sector)
mtext(side=3, line=3, at=title_at, adj=0, cex=1.3, mytitle)
mtext(side=3, line=2, at=subtitle_at, adj=0, cex=1, mysubtitle)
}
```

```{r}
Bankruptcy%>% 
 column_to_rownames('ID') %>% 
 select_if(is.numeric) %>% 
 na.omit() %>% prcomp(scale.=TRUE)->whole_pca

#screeplot
summary(whole_pca)
png(file="PCA2D_Screeplot.png", res=300, width=2000, height=1700)
screeplot(whole_pca,type = 'l')

#biplot
png(file="PCA2D_biplot.png", res=300, width=2000, height=1700)
biplot(whole_pca,cex=0.8,sub="Distance plot") 
add_title("Entire sectors", -175.1, -175.4)


```


In order to analysis the features of bankruptcy companies, we make PCA analysis in different industry. 
- In financial industry, companies are separated into three parts. The first part is center on relative higher GDP, CPI, YearField, FillingRate, such as 339 and 308. The second part is focused on relative lower employees, liab, assests, primefiling, sales, such as 270 and 173. The third part has higher headCityPop such as 63 and 46.


```{r fin-PCA}
#df1 financial
financial_df1_pca <- df1 %>% 
 column_to_rownames('ID') %>% 
 select_if(is.numeric) %>% 
 na.omit() %>% prcomp(scale.=TRUE)
  
summary(financial_df1_pca)

#Distance biplot 
png(file="PCA2D_Fin_biplot.png", res=300, width=2000, height=1700)
biplot(financial_df1_pca,sub="Distance biplot",cex=0.8) 
add_title("Financial sector",-8.1, -8.2)
```

- In property industry, companies are relative separated. No. 206 is the outlier, which is JWP Inc. It has highest sales, employees and lowest filling rate and daysin.


```{r prop-PCA}
#df2 property
property_df2_pca <- df2 %>% 
 column_to_rownames('ID') %>% 
 select_if(is.numeric) %>% 
 na.omit() %>% prcomp(scale.=TRUE)
  
summary(property_df2_pca)
#Distance biplot 
png(file="PCA2D_Prop_biplot.png", res=300, width=2000, height=1700)
biplot(property_df2_pca,sub="Distance biplot",cex=0.8) 
add_title("Property sector", -5.5,-5.6)
```

- In service industry, most companies are widely distributed. Probably those service companies do not have more similar features than other industry.


```{r serv-PCA}
service_df3_pca <- df3 %>% 
 column_to_rownames('ID') %>% 
 select_if(is.numeric) %>% 
 na.omit() %>% prcomp(scale.=TRUE)
  
summary(service_df3_pca)
#Distance biplot 
png(file="PCA2D_Serv_biplot.png", res=300, width=2000, height=1700)
biplot(service_df3_pca,sub="Distance biplot",cex=0.8)
add_title("Service sector", -15.5,-15.6)
```

- In industry, almost all companies are concentrated on the same position, which means that they all have similar features. It is worth noting No.45 and No.413. No.45 is LTV Corp, which have the highest micro features, such as highest employees, ebit, sales and liab.
No.413 is Owens Corning, which has the lowest population of the firms headquarters city. Because its headquarters city is Toledo ohio, which just has 270871 population in 2020.

```{r ind-PCA}
industry_df4_pca <- df4 %>% 
 column_to_rownames('ID') %>% 
 select_if(is.numeric) %>% 
 na.omit() %>% prcomp(scale.=TRUE)
  
#Distance biplot 
png(file="PCA2D_Ind_biplot.png", res=300, width=2000, height=1700)
biplot(industry_df4_pca,sub="Distance biplot",cex=0.8)
add_title("Industrial sector", -8.5,-8.6)
```

- In consumption industry, large parts concentrate on the arrow of 'HeadCityPop' which indicates that almost half headquarters of consumption companies in the cities with relative large populations and population might be a indicator for consumption companies to establish headquarters. For other major parts of consumption, they did not show a specific distribution with loading directions, further PCs may be needed.

```{r cons-PCA}
consumption_df5_pca <- df5 %>% 
 column_to_rownames('ID') %>% 
 select_if(is.numeric) %>% 
 na.omit() %>% prcomp(scale.=TRUE)
#Distance biplot 
png(file="PCA2D_Cons_biplot.png", res=300, width=2000, height=1700)
biplot(consumption_df5_pca,sub="Distance biplot",cex=0.8)
add_title("Consumption sector", -8.8,-8.9)
```

- In resource industry, companies spread dispersed in the biplot, most of them are on the positive side of PC2 while the variables of 'YearFiled', 'GDP' and 'FilingRate' cotributes to positive PC2. However, those companies only go over a bit of PC2 and none of variables could specifically explain their distribution so those 3 variables might no be able to explain those major observation well.



```{r res-PCA}
resource_df7_pca <- df7 %>% 
 column_to_rownames('ID') %>% 
 select_if(is.numeric) %>% 
 na.omit() %>% prcomp(scale.=TRUE)
  
png(file="PCA2D_Res_biplot.png", res=300, width=2000, height=1700)
biplot(resource_df7_pca,sub="distance biplot",cex=0.8)
add_title("Resource sector",-4.22,-4.24)
```

======================================== PCA3D ======================================== 




```{r screeplot}
bkc <- Bankruptcy %>% semi_join(pca_data, by = 'Ebit')
#PCA
bkc_pca<- prcomp(pca_data[1:14], center = TRUE, scale = TRUE) 
join <- bkc %>% mutate(PC1 = bkc_pca$x[,1], PC2 = bkc_pca$x[,2], PC3 = bkc_pca$x[,3])

bkc_pca_var <- tibble(PC=1:length(bkc_pca$sdev), prop.var=bkc_pca$sdev^2)
bkc_pcs <- as_tibble(bkc_pca$x[,1:3]) %>%
  mutate(company=bkc$Name)
bkc_pcs_evc <- as_tibble(bkc_pca$rotation[,1:3]) %>% 
  mutate(origin = rep(0,14), variables = colnames(pca_data)) %>% 
  mutate(PC1s = PC1*(bkc_pca_var$prop.var[1]*2),
         PC2s = PC2*(bkc_pca_var$prop.var[2]*2),
         PC3s = PC3*(bkc_pca_var$prop.var[3]*2)
         )
```
added in Appendix
```{r }

par(mar = c(1.2,1.2,1.2,1.2))
scatter3D(bkc_pcs$PC1, bkc_pcs$PC2, bkc_pcs$PC3, 
          pch = 20, theta = 150, phi = 10,
          ticktype = "detailed",
     colvar= as.integer(factor(bkc$DENYOther )),
     col = gg2.col(3),
     colkey = list(
       at = c(
         unique(as.integer(factor(bkc$DENYOther)))),
       side = 4, length = 1, width = 0.5, labels = unique(factor(bkc$DENYOther))),
      xlab = "PC1", ylab = "PC2", zlab = "PC3", main = "PCs agaisnt DENYOther", clab = "DENYOther"
     )
points3D(bkc_pcs_evc$origin, bkc_pcs_evc$origin, bkc_pcs_evc$origin, add = TRUE, col="blue", 
          colkey = FALSE, pch = 19, cex = 1)
arrows3D(bkc_pcs_evc$origin, bkc_pcs_evc$origin, bkc_pcs_evc$origin,
         bkc_pcs_evc$PC1s, bkc_pcs_evc$PC2s, bkc_pcs_evc$PC3s,
         phi = 0, theta = 90,
         lwd = 2, d = 3, clab = c("Quality", "score"), 
         main = "Loading", bty ="g", ticktype = "detailed", overlay = T,add = T)
text3D( bkc_pcs_evc$PC1s, bkc_pcs_evc$PC2s, bkc_pcs_evc$PC3s,  
       c(bkc_pcs_evc$variables),
       add=TRUE, colkey = FALSE, col = "red")
```

```{r biplot-sector, warning=F}
biplot_sector <- ggplot() + 
  geom_point(data=join, aes(x=PC2, y=PC3, col = Sector)) +

  xlab("PC2") + ylab("PC3") + 
  theme(aspect.ratio=1)  +
  theme_classic() +
  geom_segment(data=bkc_pcs_evc, aes(x=origin, xend=PC2s, y=origin, yend=PC3s), colour="black") +
  ggrepel:: geom_text_repel(data=bkc_pcs_evc, aes(x=PC2s, y=PC3s, label=variables), colour="red", nudge_y=sign(bkc_pcs_evc$PC3)*0.2)+
  labs(x = "PC2", y = "PC3", title = "Biplot", subtitle = "Variables projected onto PC2 and PC3") +
   theme( text=element_text(family="Times New Roman", face="bold", size=12))

png(file="PCA3D_biplot_Sector.png", res=300, width=2000, height=1000)
biplot_sector
```
```{r biplot-DNother, warning=F}
biplot_deny <- ggplot() + 
  geom_point(data=join, aes(x=PC2, y=PC3, col = DENYOther)) +
  xlab("PC2") + ylab("PC3") + 
  theme(aspect.ratio=1) +
  theme_classic() +
  geom_segment(data=bkc_pcs_evc, aes(x=origin, xend=PC2s, y=origin, yend=PC3s), colour="black") +
  ggrepel:: geom_text_repel(data=bkc_pcs_evc, aes(x=PC2s, y=PC3s, label=variables), colour="red", nudge_x=0.1, nudge_y=sign(bkc_pcs_evc$PC3)*0.2)+
  labs(x = "PC2", y = "PC3", title = "Biplot", subtitle = "Variables projected onto PC2 and PC3") +
   theme( text=element_text(family="Times New Roman", face="bold", size=12))

png(file="PCA3D_biplot_DENYother.png", res=300, width=2000, height=1000)
biplot_deny

```